```{r Delirium-dementia, include=FALSE}
# getResDelAphasia <- function(df_adm_f = df_adm_f, data_dic = data_dic){
        df <- df_adm_f
        df$var_prev_dementia <- as.factor(df$var_prev_dementia)
        dd <- datadist(df)
        options(datadist = 'dd')
        options(contrasts=c("contr.treatment", "contr.treatment"))
        
        #Variables ? include dementia?
        dv = "outcome_delir"
        iv0 = "var_prev_dementia + sex"
        iv1 = "var_prev_dementia + sex + rcs(age, 3)"
        iv2 = "var_prev_dementia + sex + rcs(nihss_d0, 3)"
        iv3 = "var_prev_dementia + sex + rcs(age, 3) + rcs(nihss_d0, 3)"
        ia = "var_prev_dementia * sex"
        #Model
        # bm_delirium_var_prev_dementia <- getModel(dv, iv0, file_name = "output/models/bm_delirium_var_prev_dementia", data = df, force_overwrite = FALSE)
        bm_delirium_var_prev_dementia <- getModel(
                dv,
                iv0,
                file_name = "output/models/bm_delirium_var_prev_dementia",
                data = df,
                force_overwrite = FALSE
        )
        
        bm_delirium_var_prev_dementia_iv3 <- getModel(
                dv,
                iv3,
                file_name = "output/models/bm_delirium_var_prev_dementia_iv3",
                data = df,
                force_overwrite = FALSE
        )
        
        #Table
        bm_delirium_var_prev_dementia_tbl <- getLatexTableCH(bm_delirium_var_prev_dementia, font_size=8)
        #interaction model
        # bm_delirium_var_prev_dementia_ia <- getModel(dv, ia, file_name = "output/models/bm_delirium_var_prev_dementia_ia", data = df, force_overwrite = FALSE)
        bm_delirium_var_prev_dementia_ia <- getModel(dv, ia, file_name = "output/models/bm_delirium_var_prev_dementia_ia", data = df, force_overwrite = FALSE)

        # Interaction Table
        bm_delirium_var_prev_dementia_tbl_ia <- getLatexTableCH(bm_delirium_var_prev_dementia_ia, font_size=8, interactions_only = TRUE)
        # bm_delirium_var_prev_dementia_ia <- getModel(dv, ia, file_name = "output/models/bm_delirium_var_prev_dementia_ia", data = df, force_overwrite = FALSE, imputation = TRUE)

        # Bivar
        iv1 = paste(ia, "+ rcs(nihss_d0, 3)")
        bm_delirium_var_prev_dementia_iv1 <- getModel(dv, iv1, file_name = "output/models/bm_delirium_var_prev_dementia_iv1", data = df, force_overwrite = FALSE)
        iv2 = paste(ia, "+ rcs(age, 3)")
        bm_delirium_var_prev_dementia_iv2 <- getModel(dv, iv2, file_name = "output/models/bm_delirium_var_prev_dementia_iv2", data = df, force_overwrite = FALSE)
        
        #Prediction
        
        bm_delirium_var_prev_dementia_p <- Predict(bm_delirium_var_prev_dementia_ia, var_prev_dementia,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        bm_delirium_var_prev_dementia_footnote <- getFootnote(bm_delirium_var_prev_dementia_p)
        # Probability difference
        bm_delirium_var_prev_dementia_contrast <- contrast(bm_delirium_var_prev_dementia_ia, 
                                             list(sex="male", var_prev_dementia=c(bm_delirium_var_prev_dementia_p$var_prev_dementia)),
                                             list(sex="female", var_prev_dementia=c(bm_delirium_var_prev_dementia_p$var_prev_dementia)))
        
        bm_delirium_var_prev_dementia_contrast_df <- data.frame(var_prev_dementia=bm_delirium_var_prev_dementia_contrast$var_prev_dementia, probability=bm_delirium_var_prev_dementia_contrast$PP)
        
        # Plots
        
        # lollipop
        bm_delirium_var_prev_dementia_p0 <- plot_grouped_bar_chart(df, "var_prev_dementia", "sex", chart_type = "lollipop") + coord_cartesian(expand=TRUE, ylim=c(0,3000))
        
        # jitter
          bm_delirium_var_prev_dementia_p1 <- ggplot(bm_delirium_var_prev_dementia_p %>% as_tibble(), aes(x = var_prev_dementia, y = yhat, color = sex, group = sex)) +
    geom_jitter(position = position_dodge(width = 0.5), size = 3) +  
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +  # Adding error bars with the same jitter
    geom_line(position = position_dodge(width = 0.5)) + 
    theme_ch() +  scale_color_colorblind()  +
    # labs(x = "Neglect Grade", y = "Predicted Value (yhat)", color = "Sex") +  # Adding labels + coor_flip()
ylab(expression(P(Delirium ~ "|" ~ Data, x)))  +
                theme(
                        # legend.position = c(0.95, 0.05),
                        # legend.justification = c(1, 0),
                        legend.position = "none",
                        legend.title = element_blank(),
                        # plot.title = element_text(hjust = 0.5),
                        panel.grid.major.x = element_blank(),
                        panel.grid.minor.x = element_blank(),
                        axis.ticks = element_line(),
                        axis.text.x = element_text(angle = 0),
                        text = element_text(size = 10)
                )
        
        bm_delirium_var_prev_dementia_p_diff <-
                plotSexDiffGradient(
                        df,
                        bm_delirium_var_prev_dementia_ia,
                        var = "var_prev_dementia",
                        xlab = expression(P(Delta ~ "|" ~ data, x) == P(theta[F]-theta[M]))) +
                coord_cartesian(expand = TRUE, ylim = c(0, 50)) +
                coord_flip()

        # 
        # bm_delirium_var_prev_dementia_p2_footnote <- paste0(expression(P(Delta ~ "|" ~ data, x)), " is the posterior probability density of the contrast between females and males of developing a delirium at a given level of dementia x.")
        # 
        library(gridExtra)
        library(ormPlot)
        library(cowplot)
        # bm_delirium_var_prev_dementia_p1 <- ggplot(Predict(bm_delirium_var_prev_dementia_ia, var_prev_dementia)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        bm_delirium_var_prev_dementia_p2 <- ggplot(Predict(bm_delirium_var_prev_dementia_ia, sex)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + coord_cartesian(expand = TRUE, ylim = c(-4, 0)) + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        bm_delirium_var_prev_dementia_p3 <- ggplot(Predict(bm_delirium_var_prev_dementia_ia, var_prev_dementia, sex, conf.int = FALSE)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D")+ theme(plot.margin = unit(c(0.5,0,0,0.5), "cm")) 
        bm_delirium_var_prev_dementia_p4 <- plot(bm_delirium_var_prev_dementia_ia) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + coord_cartesian(expand = TRUE, xlim = c(-3, 3)) + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        bm_delirium_var_prev_dementia_p5 <- pp_check.blrm(bm_delirium_var_prev_dementia_ia, type = "violin_grouped", group = "sex", df=df) + coord_cartesian(expand = TRUE) + theme(legend.position = "top", plot.margin = unit(c(0.5,0,0,1), "cm"))
        #forest plot
        # p6_temp <- forestplot(summary(bm_delirium_var_prev_dementia_ia), return_ggplots = TRUE)
        bm_delirium_var_prev_dementia_p6 <- forestplot(summary(bm_delirium_var_prev_dementia), return_ggplots = TRUE)
        bm_delirium_var_prev_dementia_p7 <- stanDxplot(bm_delirium_var_prev_dementia_ia, which="sex=male") + theme_ch() + coord_cartesian(expand = TRUE, xlim=c(0,1400), ylim=c(-10,25)) + coord_flip() + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        
        res_del_dementia <-
                list(
                        dv = dv,
                        iv0 = iv0,
                        ia = ia,
                        model = bm_delirium_var_prev_dementia,
                        modelajd = bm_delirium_var_prev_dementia_iv3,
                        tbl = bm_delirium_var_prev_dementia_tbl,
                        model_ia = bm_delirium_var_prev_dementia_ia,
                        tbl_ia = bm_delirium_var_prev_dementia_tbl_ia,
                        pred = bm_delirium_var_prev_dementia_p,
                        pred_fn = bm_delirium_var_prev_dementia_footnote,
                        p0 = bm_delirium_var_prev_dementia_p0,
                        p1 = bm_delirium_var_prev_dementia_p1,
                        p2 = bm_delirium_var_prev_dementia_p2,
                        p3 = bm_delirium_var_prev_dementia_p3,
                        p4 = bm_delirium_var_prev_dementia_p4,
                        # p5 = bm_delirium_var_prev_dementia_p5,
                        p6 = bm_delirium_var_prev_dementia_p6,
                        p7 = bm_delirium_var_prev_dementia_p7,
                        p_diff = bm_delirium_var_prev_dementia_p_diff
                )
        # return(res_del_var_prev_dementia)
        
# }
        
            

```
### Dementia

This analysis investigate frequency of a known dementia at hospital presentation between the sexes and their relationship with occureence of delirium. Data completeness for dementia status was 89%. Overall 4.2% of the patients had a known dementia with a female preponderance of 5.5% as compared to males with 2.8% (odds ratio 2.05, 95% credible interval 1.54--2.72, probability of difference >99.9%, see @tbl-delir-dementia. Having a previous diagnosis of dementia increased the likelihood of developing a delirium during hospital stay (odds ratio 1.92, 95% credible interval 1.23-- 2.96, Pr(\Beta > 0) = 0.9962). However, when adjusting for potential confounders such as age, NIHSS, and pRS with an `r getMedCI(summary(res_del_dementia$model), var="var_prev_dementia")`,  `r getP(getRsmb(res_del_dementia$model, dec=3), var="var_prev_dementia")` - this association was attenuated, mostly mediated by age.
```{r tbl-Delirium-dementia}
#| label: tbl-delir-dementia
#| tbl-cap: Descriptive analysis of _Previously Known Dementia_ in the whole cohort and subgroups of biological sex
#| tbl-colwidths: "auto"
#| echo: false
library(gt)
library(kableExtra)
# t <- getLatexTableCH(bm_delirium_age, font_size=8)
sel <- c("sex", "var_prev_dementia")
var <- setdiff(sel, "sex")
# file storage
file <- "output/models/bm_delirium_var_prev_dementia_uni.rds"
# Run Bayesian Univariate calculations
if (file.exists(file)) {
        models <- readRDS(file)
} else {
        models <- run_models(df, var)
        saveRDS(models,
                file,
                compress = "xz")
}
stat <- extract_bayes_stat(models, var)

plot_tbl(
    data = df,
    selected_columns = c("var_prev_dementia","sex"),
    by = "sex",
    stat_descr_loc = "bottom",
    pkg = "kableExtra",
    kable_option = "latex",
    pvalue = FALSE,
    bayes = TRUE,
    stat, # adding 'stat# to add bayesian stats
    missing = "ifany"
) %>% 
    kable_styling(latex_options = c("scale_down"))  
```
While females with no dementia are at a lower risk of developing a delirium compared to males, a previously known dementia in females lead to a increase in risk for delirium @fig-delirium-dementia -B (probability of interaction 93.8%). This finding was independent of age and NIHSS on delirium.
```{r fig-delirium-dementia, fig.width=6, fig.asp=0.5, out.width="100%"}
#| label: fig-delirium-dementia
#| echo: false
#| warning: false
#| error: false
#| fig-cap: !expr 'paste0("**(A)** Descriptive Analysis of _Previously Known Dementia_ stratified by biological sex. **(B)** Probability of Occurrence of a New Delirium given the presence of dementia in feales and males, **(C)** Differences in posterior probability of females - males depending on the status of dementia")'
#| cache: TRUE
p <-       plot_grid(
        res_del_dementia$p0 + theme(legend.position = c(0.83,0.95), 
                                   axis.text.x = element_text(angle = 45, hjust = 1),
                                   legend.background = element_rect(fill = NA),
                                   legend.key.size = unit(0.2, "cm")),
        res_del_dementia$p1 + ylim(0,0.3) + theme(legend.position = "", axis.text.x = element_text(angle = 45, hjust = 1)),
        res_del_dementia$p_diff + theme(legend.position = "", axis.text.x = element_text(angle = 45, hjust = 1)),
        nrow = 1,
        labels = c("A", "B", "C")
)
p
```
```{r tbl2-delirium-dementia}
#| label: tbl2-delir-dementia
#| tbl-cap: Relationship between _previously known dementia_ and occurrence of *_New Delirium_*
#| tbl-colwidths: "auto"
#| results: hold
#| echo: false
#| tbl-pos: H
library(gt)
library(kableExtra)
# t <- getLatexTableCH(bm_delirium_age, font_size=8)
res_del_dementia$tbl 
```






