### Aphasia, Sex and Delirium 
```{r Delirium-Aphasia, include=FALSE}
# getResDelAphasia <- function(df_adm_f = df_adm_f, data_dic = data_dic){
       
        
        df <- df_adm_f
        dd <- datadist(df)
        options(datadist = 'dd')
        options(contrasts=c("contr.treatment", "contr.treatment"))
        
        #Variables ? include handedness?
        dv = "outcome_delir"
        iv0 = "var_aphasia_grade2 + sex"
        ia = "var_aphasia_grade2 * sex"
        
        # MODEL
        bm_delirium_var_aphasia_grade2 <- getModel(dv, iv0, file_name = "output/models/bm_delirium_var_aphasia_grade2", data = df, force_overwrite = FALSE)
        #Table
        bm_delirium_var_aphasia_grade2_tbl <- getLatexTableCH(bm_delirium_var_aphasia_grade2, font_size=8, output_format="latex")
        
        #interaction model
        bm_delirium_var_aphasia_grade2_ia <- getModel(dv, ia, file_name = "output/models/bm_delirium_var_aphasia_grade2_ia", data = df, force_overwrite = FALSE)
        # Interaction Table
        bm_delirium_var_aphasia_grade2_tbl_ia <- getLatexTableCH(bm_delirium_var_aphasia_grade2_ia, font_size=8, interactions_only = TRUE)
        
        #Prediction
        bm_delirium_var_aphasia_grade2_p <- Predict(bm_delirium_var_aphasia_grade2_ia, var_aphasia_grade2,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        bm_delirium_var_aphasia_grade2_footnote <- getFootnote(bm_delirium_var_aphasia_grade2_p)
        
        # Probability difference
        bm_delirium_var_aphasia_grade2_contrast <- contrast(bm_delirium_var_aphasia_grade2_ia, 
                                             list(sex="male", var_aphasia_grade2=c(bm_delirium_var_aphasia_grade2_p$var_aphasia_grade2)),
                                             list(sex="female", var_aphasia_grade2=c(bm_delirium_var_aphasia_grade2_p$var_aphasia_grade2)))
        
        bm_delirium_var_aphasia_grade2_contrast_df <- 
                data.frame(var_aphasia_grade2=bm_delirium_var_aphasia_grade2_contrast$var_aphasia_grade2, probability=bm_delirium_var_aphasia_grade2_contrast$PP)
        
        # BIVARIABLE
        iv1 = paste0(ia, "+rcs(nihss_d0, 3)")
        bm_delirium_var_aphasia_grade2_iv1 <- getModel(dv, iv1, file_name = "output/models/bm_delirium_var_aphasia_grade2_iv1", data = df, force_overwrite = FALSE)
        
        # bm_delirium_var_aphasia_grade2_p2 <- Predict(bm_delirium_var_aphasia_grade2_iv1, nihss_d0, var_aphasia_grade2,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        bm_delirium_var_aphasia_grade2_iv1_pred2 <- Predict(bm_delirium_var_aphasia_grade2_iv1, nihss_d0=12, var_aphasia_grade2,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        
        bm_delirium_var_aphasia_grade2_footnote2 <- getFootnote(bm_delirium_var_aphasia_grade2_iv1_pred2)
        
        
        
        # PLOTS 
        # lollipop
        bm_delirium_var_aphasia_grade2_p0 <- plot_grouped_bar_chart(df, "var_aphasia_grade2", "sex", chart_type = "lollipop") 
        
          bm_delirium_var_aphasia_grade2_p1 <- plotInteraction(bm_delirium_var_aphasia_grade2_iv1,
                                                               bm_delirium_var_aphasia_grade2_iv1_pred2,
                                                               var_of_interest = "var_aphasia_grade2",
                                                               group_var = "sex",
                                                               display_annotations = FALSE,
                                                               display_sign = TRUE)
          
        
        bm_delirium_var_aphasia_grade2_p_diff <-
                plotSexDiffGradient(
                        df,
                        bm_delirium_var_aphasia_grade2_iv1,
                        var = "var_aphasia_grade2",
                        xlab = expression(P(Delta ~ "|" ~ data, x) == P(theta[F]-theta[M]))) +
                coord_flip()
        

        bm_delirium_var_aphasia_grade2_p2_footnote <- paste0(expression(P(Delta ~ "|" ~ data, x)), " is the posterior probability density of the contrast between females and males of developing a delirium at a given level of aphasia x.")
        
        library(gridExtra)
        library(ormPlot)
        library(cowplot)
        # bm_delirium_var_aphasia_grade2_p1 <- ggplot(Predict(bm_delirium_var_aphasia_grade2_ia, var_aphasia_grade2)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        bm_delirium_var_aphasia_grade2_p2 <- 
                ggplot(Predict(bm_delirium_var_aphasia_grade2_ia, sex)) + 
                theme_ch() + 
                scale_color_viridis(discrete = TRUE, option = "D") + 
                coord_cartesian(expand = TRUE, ylim = c(-4, 0)) + 
                theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        bm_delirium_var_aphasia_grade2_p3 <- 
                ggplot(Predict(bm_delirium_var_aphasia_grade2_ia, var_aphasia_grade2, sex, conf.int = FALSE)) + 
                theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm")) 
        
        bm_delirium_var_aphasia_grade2_p4 <-
                plot(bm_delirium_var_aphasia_grade2_ia) + 
                theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + 
                coord_cartesian(expand = TRUE, xlim = c(-3, 3)) + 
                theme(plot.margin = unit(c(0.5, 0, 0, 0.5), "cm"))
        
        bm_delirium_var_aphasia_grade2_p5 <- 
                pp_check.blrm(bm_delirium_var_aphasia_grade2_ia, type = "violin_grouped", group = "sex", df=df) + 
                coord_cartesian(expand = TRUE) + 
                theme(legend.position = "top", plot.margin = unit(c(0.5,0,0,1), "cm"))
       
         #forest plot
        # p6_temp <- forestplot(summary(bm_delirium_var_aphasia_grade2_ia), return_ggplots = TRUE)
        bm_delirium_var_aphasia_grade2_p6 <- forestplot(summary(bm_delirium_var_aphasia_grade2_ia), return_ggplots = TRUE)
        bm_delirium_var_aphasia_grade2_p7 <- 
                stanDxplot(bm_delirium_var_aphasia_grade2_ia, which="sex=male") + 
                theme_ch() + coord_cartesian(expand = TRUE, xlim=c(0,1400), ylim=c(-10,25)) + 
                coord_flip() + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        
        res_del_aphasia <-
                list(
                        dv = dv,
                        iv0 = iv0,
                        ia = ia,
                        model = bm_delirium_var_aphasia_grade2,
                        tbl = bm_delirium_var_aphasia_grade2_tbl,
                        model_ia = bm_delirium_var_aphasia_grade2_ia,
                        tbl_ia = bm_delirium_var_aphasia_grade2_tbl_ia,
                        pred = bm_delirium_var_aphasia_grade2_iv1_pred2,
                        pred_fn = bm_delirium_var_aphasia_grade2_footnote,
                        p0 = bm_delirium_var_aphasia_grade2_p0,
                        p1 = bm_delirium_var_aphasia_grade2_p1,
                        p2 = bm_delirium_var_aphasia_grade2_p2,
                        p3 = bm_delirium_var_aphasia_grade2_p3,
                        p4 = bm_delirium_var_aphasia_grade2_p4,
                        p5 = bm_delirium_var_aphasia_grade2_p5,
                        p6 = bm_delirium_var_aphasia_grade2_p6,
                        p7 = bm_delirium_var_aphasia_grade2_p7,
                        p_diff = bm_delirium_var_aphasia_grade2_p_diff
                )
        # return(res_del_var_aphasia_grade2)
        
# }
        
        most_category_aphasia <- df %>%
                filter(var_aphasia_grade2 != "no aphasia") %>%
                group_by(var_aphasia_grade2) %>%
                count() %>% ungroup() %>%
                filter(n == max(n))
        
        moderate_P = getP(getRsmb(bm_delirium_var_aphasia_grade2_ia), var = "var_aphasia_grade2=moderate * sex=male")

```
Symptoms of aphasia of any severity grade were found in n = `r nrow(df[df$var_aphasia_grade2 != "no aphasia", ])` (`r round(nrow(df[df$var_aphasia_grade2 != "no aphasia", ])/nrow(df)*100,1)`%) patient with available data (missing n = `r nrow(df[is.na(df$var_aphasia_grade2), ])`). More female patients than males presented with symptoms of aphasia. Aphasia with the grade `r most_category_aphasia[[1]]` was the most frequently occurring, affecting 29.9% of the patients, with a slightly higher occurrence in females (31.5%) compared to males (28.4%). (@tbl-delir-aphasia)
```{r tbl-Delirium-Aphasia}
#| label: tbl-delir-aphasia
#| tbl-cap: Descriptive analysis of the *degree of aphasic symptoms* in the whole cohort and subgroups of biological sex
#| tbl-colwidths: "auto"
#| echo: false
library(gt)
library(kableExtra)
# t <- getLatexTableCH(bm_delirium_age, font_size=8)
bm <- blrm(var_aphasia_grade2~sex, data = df)
stat <- getRsmb(bm, col_var=TRUE) %>% as_tibble() %>% rename(`P(β>0)`=P)

plot_tbl(
    data = df,
    selected_columns = c("var_aphasia_grade2","sex"),
    by = "sex",
    stat_descr_loc = "bottom",
    pkg = "kableExtra",
    kable_option = "latex",
    pvalue = FALSE,
    bayes = TRUE,
    missing = "ifany",
    stat # adding 'stat# to add bayesian stats
) %>% 
    kable_styling(latex_options = c("scale_down"))  




```
The bayesian regression analysis investigates sex differences in the relationship between _levels of aphasia_ and the _risk of delirium_. While aphasia severity alone (main effect) shows variable and uncertain effects on delirium, the relevance of an interaction between aphasia severity and sex was evident. Males with moderate ($P_{interaction}(\beta>0):$ `r moderate_P`) to severe aphasia ($P_{interaction} (\beta>0):$ `r getP(getRsmb(bm_delirium_var_aphasia_grade2_ia), var = "var_aphasia_grade2=severe * sex=male")`) are significantly more likely to develop delirium compared to females with the same level of aphasia. (@fig-delirium-aphasia A)
Inclusion of NIHSS (as a potential confounder of this interaction) demonstrated further strength of these findings. This suggests that males with moderate to severe aphasia are particularly vulnerable to delirium.
```{r tbl2-Delirium-Aphasia}
#| label: tbl2-delir-aphasia
#| tbl-cap: Relationship between _degree of aphasic symptoms_ and occurrence of *_New Delirium_*
#| tbl-colwidths: "auto"
#| results: hold
#| echo: false
#| tbl-pos: H
library(gt)
library(kableExtra)
res_del_aphasia$tbl
```
```{r fig-Delirium-Aphasia, fig.width=6, fig.asp=0.5, out.width="100%"}
#| label: fig-delirium-aphasia
#| echo: false
#| warning: false
#| error: false
#| fig-cap: !expr 'paste0("**(A)** Descriptive Analysis of Levels of Aphasia (no, minor, moderate, severe, global) stratified by biological sex. **(B)** Probability of Occurrence of a New Delirium given the different levels of aphasia comparing females and males, **(C)** Differences in posterior probability of females - males in each level of aphasia.")'
#| cache: TRUE
p <-       plot_grid(
        res_del_aphasia$p0 + ylim(0, 2000) + theme(legend.position = c(0.83,0.95), 
                                   axis.text.x = element_text(angle = 45, hjust = 1),
                                   legend.background = element_rect(fill = NA),
                                   legend.key.size = unit(0.2, "cm")),
        res_del_aphasia$p1 + ylim(0,0.3) + theme(legend.position = "", axis.text.x = element_text(angle = 45, hjust = 1)),
        res_del_aphasia$p_diff +  theme(axis.text.x = element_text(angle = 45, hjust = 1)),
        nrow = 1,
        labels = c("A", "B", "C")
)
p
```

### Neglect, Sex and Delirum
```{r Delirium-Neglect, include=FALSE}
# getResDelNeglect <- function(df_adm_f = df_adm_f, data_dic = data_dic){
       
        
        df <- df_adm_f
        dd <- datadist(df)
        options(datadist = 'dd')
        options(contrasts=c("contr.treatment", "contr.treatment"))
        
        #Variables ? include handedness?
        dv = "outcome_delir"
        iv0 = "var_neglect_grade + sex"
        ia = "var_neglect_grade * sex"
        
        #Model
        bm_delirium_var_neglect_grade <- getModel(dv, iv0, file_name = "output/models/bm_delirium_var_neglect_grade", data = df, force_overwrite = FALSE)
        #Table
        bm_delirium_var_neglect_grade_tbl <- getLatexTableCH(bm_delirium_var_neglect_grade, font_size=8)
        #interaction model
        bm_delirium_var_neglect_grade_ia <- getModel(dv, ia, file_name = "output/models/bm_delirium_var_neglect_grade_ia", data = df, force_overwrite = FALSE)
        # Interaction Table
        bm_delirium_var_neglect_grade_tbl_ia <- getLatexTableCH(bm_delirium_var_neglect_grade_ia, font_size=8, interactions_only = TRUE)
        
         # BIVARIABLE
        iv1 = paste0(ia, "+rcs(nihss_d0, 3)")
        bm_delirium_var_neglect_grade_iv1 <- getModel(dv, iv1, file_name = "output/models/bm_delirium_var_neglect_grade_iv1", data = df, force_overwrite = FALSE)
       bm_delirium_var_neglect_grade_iv1_pred2 <- Predict(bm_delirium_var_neglect_grade_iv1, nihss_d0=12, var_neglect_grade,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        #Prediction
        
        bm_delirium_var_neglect_grade_p <- Predict(bm_delirium_var_neglect_grade_ia, var_neglect_grade,  sex, funint=FALSE, fun=plogis, conf.int = TRUE)
        bm_delirium_var_neglect_grade_footnote <- getFootnote(bm_delirium_var_neglect_grade_p)
        # Probability difference
        bm_delirium_var_neglect_grade_contrast <- contrast(bm_delirium_var_neglect_grade_ia, 
                                             list(sex="male", var_neglect_grade=c(bm_delirium_var_neglect_grade_p$var_neglect_grade)),
                                             list(sex="female", var_neglect_grade=c(bm_delirium_var_neglect_grade_p$var_neglect_grade)))
        
        bm_delirium_var_neglect_grade_contrast_df <- data.frame(var_neglect_grade=bm_delirium_var_neglect_grade_contrast$var_neglect_grade, probability=bm_delirium_var_neglect_grade_contrast$PP)
        
        # Plots
        #TODO  implement jitter
        # bm_delirium_var_neglect_grade_p1 <- ggplot(bm_delirium_var_neglect_grade_p,
        #                              adj.subtitle = 0) + 
        #         geom_jitter(width = 0.2, height = 0, size = 5) + 
        #         theme_minimal() + scale_color_colorblind()  +
        #         ylab(expression(P(Delirium ~ "|" ~ Data, x)))  +
        #         theme(
        #                 # legend.position = c(0.95, 0.05),
        #                 # legend.justification = c(1, 0),
        #                 legend.position = "bottom",
        #                 legend.title = element_blank(),
        #                 # plot.title = element_text(hjust = 0.5),
        #                 panel.grid.major.x = element_blank(),
        #                 panel.grid.minor.x = element_blank(),
        #                 axis.ticks = element_line(),
        #                 axis.text.x = element_text(angle = 0),
        #                 text = element_text(size = 10)
        #         )
        # ) +
        # scale_x_continuous(sec.axis = sec_axis(~ . , name = "", breaks = NULL, labels = NULL))
        
        # jitter
        
#         bm_delirium_var_neglect_grade_p1 <- ggplot(bm_delirium_var_neglect_grade_p %>% as_tibble(), aes(x = var_neglect_grade, y = yhat, color = sex, group = sex)) +
#     geom_jitter(position = position_dodge(width = 0.5), size = 3) +  # Adding jittered points
#     geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +  # Adding error bars with the same jitter
#     geom_line(position = position_dodge(width = 0.5)) +  # Adding a line connecting the points
#     theme_ch() +  scale_color_colorblind()  +
#     labs(x = "Neglect Grade", y = "Predicted Value (yhat)", color = "Sex") +  # Adding labels + coor_flip()
# ylab(expression(P(Delirium ~ "|" ~ Data, x)))  +
#                 theme(
#                         # legend.position = c(0.95, 0.05),
#                         # legend.justification = c(1, 0),
#                         legend.position = "bottom",
#                         legend.title = element_blank(),
#                         # plot.title = element_text(hjust = 0.5),
#                         panel.grid.major.x = element_blank(),
#                         panel.grid.minor.x = element_blank(),
#                         axis.ticks = element_line(),
#                         axis.text.x = element_text(angle = 0),
#                         text = element_text(size = 10)
#                 )
        
          bm_delirium_var_neglect_grade_p0 <- plot_grouped_bar_chart(df, "var_neglect_grade", "sex", chart_type = "lollipop") 
       
        bm_delirium_var_neglect_grade_p1 <- plotInteraction(bm_delirium_var_neglect_grade_iv1,
                                                               bm_delirium_var_neglect_grade_iv1_pred2,
                                                               var_of_interest = "var_neglect_grade",
                                                               group_var = "sex",
                                                               display_annotations = TRUE, display_sign=FALSE)
        
        
        
        bm_delirium_var_neglect_grade_p_diff <-
                plotSexDiffGradient(
                        df,
                        bm_delirium_var_neglect_grade_ia,
                        var = "var_neglect_grade",
                        xlab = expression(P(Delta ~ "|" ~ data, x) == P(theta[F]-theta[M]))) +
                coord_flip()
        
        bm_delirium_var_neglect_grade_p2_footnote <- paste0(expression(P(Delta ~ "|" ~ data, x)), " is the posterior probability density of the contrast between females and males of developing a delirium at a given level of neglect x.")
        
        library(gridExtra)
        library(ormPlot)
        library(cowplot)
        # bm_delirium_var_neglect_grade_p1 <- ggplot(Predict(bm_delirium_var_neglect_grade_ia, var_neglect_grade)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        bm_delirium_var_neglect_grade_p2 <- ggplot(Predict(bm_delirium_var_neglect_grade_ia, sex)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + coord_cartesian(expand = TRUE, ylim = c(-4, 0)) + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        bm_delirium_var_neglect_grade_p3 <- ggplot(Predict(bm_delirium_var_neglect_grade_ia, var_neglect_grade, sex, conf.int = FALSE)) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D")+ theme(plot.margin = unit(c(0.5,0,0,0.5), "cm")) 
        bm_delirium_var_neglect_grade_p4 <- plot(bm_delirium_var_neglect_grade_ia) + theme_ch() + scale_color_viridis(discrete = TRUE, option = "D") + coord_cartesian(expand = TRUE, xlim = c(-3, 3)) + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        bm_delirium_var_neglect_grade_p5 <- pp_check.blrm(bm_delirium_var_neglect_grade_ia, type = "violin_grouped", group = "sex", df=df) + coord_cartesian(expand = TRUE) + theme(legend.position = "top", plot.margin = unit(c(0.5,0,0,1), "cm"))
        #forest plot
        # p6_temp <- forestplot(summary(bm_delirium_var_neglect_grade_ia), return_ggplots = TRUE)
        bm_delirium_var_neglect_grade_p6 <- forestplot(summary(bm_delirium_var_neglect_grade), return_ggplots = TRUE)
        bm_delirium_var_neglect_grade_p7 <- stanDxplot(bm_delirium_var_neglect_grade_ia, which="sex=male") + theme_ch() + coord_cartesian(expand = TRUE, xlim=c(0,1400), ylim=c(-10,25)) + coord_flip() + theme(plot.margin = unit(c(0.5,0,0,0.5), "cm"))
        
        res_del_neglect <-
                list(
                        dv = dv,
                        iv0 = iv0,
                        ia = ia,
                        model = bm_delirium_var_neglect_grade,
                        tbl = bm_delirium_var_neglect_grade_tbl,
                        model_ia = bm_delirium_var_neglect_grade_ia,
                        tbl_ia = bm_delirium_var_neglect_grade_tbl_ia,
                        pred = bm_delirium_var_neglect_grade_p,
                        pred_fn = bm_delirium_var_neglect_grade_footnote,
                        p0 = bm_delirium_var_neglect_grade_p0,
                        p1 = bm_delirium_var_neglect_grade_p1,
                        p2 = bm_delirium_var_neglect_grade_p2,
                        p3 = bm_delirium_var_neglect_grade_p3,
                        p4 = bm_delirium_var_neglect_grade_p4,
                        p5 = bm_delirium_var_neglect_grade_p5,
                        p6 = bm_delirium_var_neglect_grade_p6,
                        p7 = bm_delirium_var_neglect_grade_p7,
                        p_diff = bm_delirium_var_neglect_grade_p_diff
                )
        # return(res_del_var_neglect_grade)
        
# }
        
        most_category_neglect <- df %>%
                filter(var_neglect_grade != "no neglect") %>%
                group_by(var_neglect_grade) %>%
                count() %>% ungroup() %>%
                filter(n == max(n))
        
        

```
Symptoms of neglect of several severity grades were found in n = `r nrow(df[df$var_neglect_grade != "no aphasia", ])` (`r round(nrow(df[df$var_neglect_grade != "no neglect", ])/nrow(df)*100,1)`%) patient (missing n = `r nrow(df[is.na(df$var_neglect_grade), ])`). The frequencies in the different levels of neglect syndroms were quite balanced betwenn females and males. (@tbl-delir-neglect) Neglect with the grade `r most_category_neglect[[1]]` was the most frequently occurring among those with neglect, affecting 16.0% of the patients, with a slightly higher occurrence in females (16.9%) compared to males (15.1%). (@tbl-delir-neglect)
```{r tbl-Delirium-Neglect}
#| label: tbl-delir-neglect
#| tbl-cap: Descriptive Analsis of _Neglect-Syndromes_ in the whole cohort and subgroups of biological sex 
#| tbl-colwidths: "auto"
#| echo: false
library(gt)
library(kableExtra)
# t <- getLatexTableCH(bm_delirium_age, font_size=8)
bm <- blrm(var_neglect_grade~sex, data = df)
stat <- getRsmb(bm, col_var=TRUE) %>% as_tibble() %>% rename(`P(β>0)`=P)

plot_tbl(
    data = df,
    selected_columns = c("var_neglect_grade","sex"),
    by = "sex",
    stat_descr_loc = "bottom",
    pkg = "kableExtra",
    kable_option = "latex",
    pvalue = FALSE,
    bayes = TRUE,
    missing = "ifany",
    stat # adding 'stat# to add bayesian stats
) %>% 
    kable_styling(latex_options = c("scale_down"))
```
The bayesian regression analysis investigates sex differences in the relationship between levels of neglect and the risk of delirium. Compared to having no neglect, minor, visual, tactile, and multimodal neglect showed high probability of an positive association with occurrence of new delirium with  $P(\beta > 0)$ of >99%, 98.0%, 89.7%, and 99.9%, respectively. (@tbl-delir-neglect2) 
```{r fig-Delirium-Neglect, fig.width=6, fig.asp=0.5, out.width="100%"}
#| label: fig-delirium-neglect
#| echo: false
#| warning: false
#| error: false
#| fig-cap: !expr 'paste0("**(A)** Descriptive Analysis of Levels of Neglect (no neglect < minor/unspecified < visual < sensible/tactile < multimodal) stratified by biological sex. **(B)** Probability of Occurrence of a New Delirium given the different levels of neglect comparing females and males, **(C)** Differences in posterior probability of females - males in each level of aphasia.")'
#| cache: TRUE
p <-       plot_grid(
        res_del_neglect$p0 + ylim(0, 2500) + theme(legend.position = c(0.83,0.95), 
                                   axis.text.x = element_text(angle = 45, hjust = 1),
                                   legend.background = element_rect(fill = NA),
                                   legend.key.size = unit(0.2, "cm")),
        res_del_neglect$p1 + theme(legend.position = "", axis.text.x = element_text(angle = 45, hjust = 1)),
        res_del_neglect$p_diff + theme(legend.position = "", axis.text.x = element_text(angle = 45, hjust = 1)),
        nrow = 1,
        labels = c("A", "B", "C")
)
p
```
Visual neglect showed a moderation effect by sex ($P_{interaction}(\beta>0)$: 0.013). That is, males showed a reduced likelihood of a new delirium compared to females in this context. Allowing for stroke severity to enter the model, the interaction of visual neglect with sex was confirmative, with a female association for increased risk of new delirium; in all other categories, males were at increased delirium risk. @fig-delirium-neglect
```{r tbl-Delirium-Neglect2}
#| label: tbl-delir-neglect2
#| tbl-cap: Relationship between _Neglect_ syndromes and occurrence of *_New Delirium_* in groups of biological sex
#| tbl-colwidths: "auto"
#| echo: false
replace_H_with(res_del_neglect$tbl)
```


